Learning CNN filters from user-drawn image markers for coconut-tree image classification

### 2020-07-24 IEEE-GRSL-Response-Letter.tex

Dear Editor, please find attached the revised version of our manuscript, GRSL-00821-2020. We thank the Editor and
reviewers for the opportunity and comments to improve the presentation of our work. We have addressed all raised issues and highlighted the changes in red. Our point-by-point response to the reviewers' comments follows next.

\rwr{1} 
The manuscript entitled "Learning CNN ﬁlters from user-drawn image-makers for coconut-tree image classiﬁcation" presents a method that needs a minimal set of user-selected images to train the CNN’s feature extractor, reducing the number of required images to train the fully connected layers. The theme of this manuscript is interesting, and the text brings an important contribution to the remote sensing area using deep learning tools. However, to be considered adequated to publish in the journal "Geoscience and Remote Sensing Letters", it requires a revision on the text. The attached document presents all corrections and observations made for it. However, the inclusion of more experiments using more images to train the proposed model is a crucial issue in my point of view. This is important to prove the robustness of the proposed model.

\ans

Thanks for the comments. We have improved the experimental section with new experiments, demonstrating the robustness of the proposed feature learning approach. We have also revised the text, added a figure, and clarified some parts to improve readability. All changes are in red.

@@

\eans

\rwr{1} A well-trained CNN model should detect image regions that best discriminate classes as positive activations, improve class separation by eliminating negative activations, and aggregate the resulting activations locally (pooling) and globally (flattening) into a highly dimensional and sparse feature space, in which a linear classifier should separate the classes with reasonable success.

\ans  Thanks. The sentence was long and unclear. We have simplified and improved it in the first paragraph of Section II.
\eans

@@

\rwr{1} with resolution under 10 $cm^2$; LAB color space; 11387 regions; MLP classifier;

\ans Thanks. These are minor corrections in different parts of Section III that have been solved now. \eans

@@

\rwr{1} We repeated this procedure three times to compute the mean and the standard deviation of the results in precision, recall, and f-score.

\ans 
This sentence was removed. The dataset was randomly divided into training, validation, and testing sets. The validation set was used only once to choose empirically the architecture of the CNN through some preliminary tests. The mean results with standard deviation were computed based on those three random splits, using their respective training and testing sets. We hope this is clear in Section III now. 
\eans

@@

\rwr{1} we identified cohesive groups of images from each class and selected two representative images from each class to draw markers.}

\ans We have changed the way to explain the choice of representative images from each class to draw markers. The new Figure 3 also helps in that explanation. \eans

@@

\rwr{1} We have not investigated the impact of placing markers on a higher number of images and network architecture optimization procedures yet.

\ans Thanks a lot for the suggestion related to this sentence. The sentence has been removed and we have added ex\eans

### 2020-07-24 main.tex
Updates from Overleaf
update model figure

\begin{figure*}
  \begin{center}
  \tikzset{
    treenode/.style = {shape=rectangle, rounded corners,
                       draw=black, align=center, line width=width, minimum width=2cm},
    specialedge/.style = {line width=1.5pt}
  }

  \begin{tikzpicture}
    [
      grow                    = right,
      sibling distance        = 5em,
      level distance          = 30mm,
      edge from parent/.style = {draw=black, -latex},
      every node/.style       = {treenode},
      sloped
    ]
    \node [treenode, fill=tealDeer] {MarkerBased \\ BatchNorm}
    child {
        node [treenode, fill=tealDeer] {Conv$(60 \times 7 \times 7)$}
        child {
          node [treenode, fill=tealDeer] {ReLU} 
          child {
            node [treenode, fill=tealDeer] {MaxPool$(3\times3)$}
            child {
              node [treenode, fill=tealDeer] {BatchNorm}
              child {
                node [treenode, fill=navyPurple] {Classifier}
                edge from parent
              }
            }
          } 
        }
      };

@@

    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
    execute at end node={\end{varwidth}}]
    \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]

    \tikzstyle{myarrow} = [navy, thick]

@@

\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set. The validation set was used to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}. The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}. It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$. Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}. We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}). Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}). These images must come from regions in the projection populated by samples from the same class (either red or green points). Those regions indicate groups of images from the same class in the input feature space. The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}

>>>>

    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=dodgerBlue, very thick, fill=dodgerBlue!25, execute at begin node={\begin{varwidth}{2.5cm}},
      execute at end node={\end{varwidth}}]

      \tikzstyle{myarrow} = [navy, very thick]

  \begin{tikzpicture}

    \node (mbn) [steps] {MarkerBased \\ BatchNorm};

    \node (conv) [steps, right= of mbn] {Conv$(60 \times 7 \times 7)$};

    \node (relu) [steps, right= of conv] {ReLU};

    \node (pool) [steps, right=of relu] {MaxPool$(3\times3)$};

    \node (bn) [steps, right=of pool] {BatchNorm};

    \node (clf) [steps, fill=navyPurple!25, draw=navyPurple, right= of bn] {Classifier};

    %-----------------------arrows---------------

    \draw [arrow, myarrow] (mbn) -- (conv);
    \draw [arrow, myarrow] (conv) -- (relu);
    \draw [arrow, myarrow] (relu) -- (pool);
    \draw [arrow, myarrow] (pool) -- (bn);
    \draw [arrow, myarrow] (bn) -- (clf);

@@

    \tikzstyle{steps} = [rectangle, rounded corners, minimum width=1.3cm, minimum height=0.5cm,text centered, draw=pink, very thick, fill=heliotrope!30, execute at begin node={\begin{varwidth}{2.5cm}},
    execute at end node={\end{varwidth}}]
    \tikzstyle{image} = [minimum width=1.3cm, minimum height=0.7cm,text centered, draw=white,fill=white!30, inner sep=0pt]

    \tikzstyle{myarrow} = [navy, very thick]

@@

\textcolor{red}{First, we evaluated H1 and H2 by randomly selecting 200 images for the training set, 1000 images for a validation set, and the remaining 12387 images were left in the testing set. The validation set was used only once to choose empirically the architecture of the network shown in Figure~\ref{fig:arch}. The convolutional layer has $60$ filters with dimension $7 \times 7$ and 3 color \textcolor{red}{bands}. It includes marker-based normalization before convolution, ReLu activation, and a max-pooling operation with a window of dimension $3 \times 3$. Since it is the single and last convolutional layer, we applied a stride of $4$ in max-pooling and batch normalization to create an input as close as possible to the one of the \textcolor{red}{MLP} classifier used in the popular VGG-16~\cite{simonyan2014very}. We then vectorized and projected the training images by t-SNE~\cite{maaten2008visualizing} (Figure~\ref{fig:pipeline-xtractor}). Four representative images, only two per class, were chosen by the user to draw markers (Figure~\ref{fig:markers}). These images must come from regions in the projection populated by samples from the same class (either red or green points). Those regions indicate groups of images from the same class in the input feature space. The markers must be placed in regions that best distinguish the coconut and non-coconut classes.}
